### 2022-07-24 main/paper3.tex
Add files via upload

\section{Introduction}
\label{S:1}

Recent developments in artificial intelligence and the growing amount of alternative data have created new areas of research in finance. In particular, often coming from news, social media or annual reports, textual data is increasingly used in the literature. \citet{nikfarjam2010text} documents recent text mining approaches for stock market prediction. The pioneer work from \citet{antweiler2004all} computes a bullishness measure out of 1.5 million messages posted on Yahoo! Finance and Raging Bull and finds that stock messages help predict market volatility. Their results clearly reject the hypothesis that all that talk is just noise. They show that there is financially relevant information present. Growing from that, there are now two main sides of sentiment analysis literature studying the forecasting abilities of natural language processing (NLP hereafter): text mining of annual reports and sentiment analysis using social media. On one side, \citet{shirata2011extracting} and \citet{cecchini2010making} report that extracting phrases from annual reports may be an effective predictor of corporate bankruptcy. Similarly, \citet{loughran2011barron} identify phrases that might be red flags indicating questionable behavior. On the other side, \citet{renault2017intraday} builds an intraday investor sentiment indicator using messages and finds that the change in investor sentiment of the first half-hour of a trading day helps forecast the last half-hour market return of that trading day.

We conjecture that our investor sentiment measure can serve as a proxy of unobservable firm fundamentals. For instance, misaligned managerial and shareholder incentives are not easy to observe quantitatively in firm's annual reports (see \citet{nikolov2014agency}) but analysts may talk about them freely in bearish messages. This flow of data can be used throughout a period to improve forecasts (see \textit{nowcasting} in \citet{challet2013predicting}). \citet{tetlock2008more} uses Wall Street Journal stories to examine whether the usage of language is able to predict individual firms’ accounting earnings and stock returns. They find that some aspects of firms' fundamentals are hard-to-quantify, but investors may use linguistic media content to capture information and incorporate it into stock prices.

Most papers studying social media's predictive power use Twitter as their primary source of data. Twitter has the advantage of being used by a wide range of people across the world and a few influencers can attract attention of many investors. In 2013, Carl Icahn tweeted that following a meeting with Tim Cook (Apple CEO), he bought a large position in Apple and believed that the company is extremely undervalued. This bullish tweet caused the market capitalization of Apple to jump by \$12 billion. In 2019, JPMorgan has created the \textit{Volfefe Index} to track Donald Trump's tweets impact on the stock market. However, it is more difficult to disentangle noise from relevant tweets in Twitter than in other more focused social media. Results from \citet{Ghoshal} show that StockTwits is significantly more informative than Twitter data. This is not surprising as StockTwits is a finance-only social media whereas Twitter also captures irrelevant opinions on a wide range of non-finance related matters.

To the best of our knowledge, our paper is the first that compares predictive power of messages in general and around specific events. \citet{ranco2015effects} is likely the closest paper to our study. However, the finance-tailored data we use allow us to get higher contemporaneous correlations between stock returns and polarity. We also use much more messages (90 million versus 1 million) during a longer period (10 years versus 13 months). We also add contribution by looking at the predictive power of cumulative average abnormal polarity (CAAP hereafter). Our paper also contributes to the Efficient Market Hypothesis (EMH hereafter) literature by gauging how cumulative average abnormal returns (CAAR hereafter) and CAAP behave around sudden peaks of message activity. We collect 90 million messages published on StockTwits from mid-2010 to the end of March 2020. On this microblogging platform, users are invited to identify firms with their ticker when they share opinions. Another useful feature on the platform is the possibility to explicitly label their own messages as \textit{bearish} or \textit{bullish} when users post them. We believe that messages on the platform are reliable for several reasons. First, users have incentives to publish valuable information in order to maintain or increase mentions and/or retweets and thus have a greater share of voice in the forum (\citet{sprenger2014tweets}). In addition, market manipulations happen rarely because malicious users have incentives to post fake news only if they previously traded in the same direction than the news they are creating, which will only benefit them if they already have influence on the platform. Finally, SEC closely monitors large influencers to prevent any market abuse.

The challenge in this context is to create a classifier that understands the vocabulary of the messages posted by the investors. For instance, ``bull" is an animal in everyday language but it is someone optimistic in the financial jargon. Work has already been done in this direction: \citet{loughran2011liability} creates a word list, which helps classify tone in a financial document. However, this might not be sufficient in the context of social media because messages posted present many typos, abbreviations and slang, so one needs to have an additional layer of data preprocessing. For instance, the word ``goooooood" would not be recognized by the model if it is not corrected into ``good" first.

In this paper, we use a logistic regression on Term Frequency-Inverse Document Frequency (TFIDF hereafter) vectorized labeled messages to classify unlabeled messages in either bearish, bullish or neutral class. TFIDF is a weighting scheme gauging the importance of a word in a document (see \citet{erdemlioglu2017market}). As users are prone to post more bullish messages than bearish messages, a good classifier needs to take into account the unbalanced data. Without resampling, the classifier outputs classification scores biased towards the over-represented class. Because we are creating an artificial neutral class, we chose to not resample the data but rather optimally select classification score thresholds by maximizing F1 scores of two distinct classification algorithms: bullish versus non-bullish and bearish versus non-bearish. Then, we aggregate messages daily for every firm to compute sentiment polarity time series for individual firms and for the whole economy. 
%\todo{we should find a concise definition of polarity to be put in the abstract and introduction} \TODO{agree. what about something like : degree of bullishness observed on a given firm on a given day}
%Similarly, \citet{pang2002thumbs} suggests that ``there are certain words people tend to use to express strong sentiments, so that it might suffice to simply produce a list of such words by introspection and rely on them alone to classify the texts."

We then use the daily volume of messages on a given firm to identify sudden peak of activity, indicating a firm event. Using abnormal polarity on each event date, we are able to classify events into three classes: bullish, neutral and bearish. We then compute cumulative average abnormal return and cumulative average abnormal polarity in a 41 days window centered at the identified event. We show that abnormal polarities have significantly higher predictive power than abnormal returns. On average, changes of polarity are associated with changes of contemporaneous return of the same sign; but this result does not hold against next-day return. However, when we focus on specific events, polarity has strong predictive power. It is also interesting to note that polarity tends to be biased towards recent past events. Finally, as robustness check, our event study on CAAR is similar to previous literature and is consistent with Fama’s theory.

The remainder of the paper is structured as follows. Section \ref{S:data} presents the data. Section \ref{S:classification} develops the NLP logistic classifier on TFIDF vectorization. Section \ref{S:polarity} explains our polarity measure. Section \ref{S:event studies} contains the results of the event studies. Section \ref{S:portfolios} shows the portfolio construction and Section \ref{S:conclusion} concludes.

\section{Data}
\label{S:data}

Data is coming from two sources: Compustat/CRSP for the stock prices, StockTwits for the investor sentiment. From the CRSP database, we extract daily stock prices (closing prices), daily volume of transactions and number of shares outstanding from 2010 to 2020 for all US and Canadian firms. Stock prices and number of shares are adjusted to account for any distribution (i.e. dividends, stock splits) so that a comparison can be made on an equivalent basis before and after the distribution.

StockTwits is a large social network similar to Twitter but designed for investors and traders. Users register online and can post messages about any listed firm through the prefix \$ followed by the ticker of the firm. StockTwits was created in 2008 as an app built on the Twitter's API and later detached from Twitter to build a standalone social network. As of April 2019, it has over two million registered users and the number of messages posted is growing exponentially (see Figure \ref{fig:tweet_vol_daily}). StockTwits describes itself as ``the voice of social finance and the best way to find out what is happening right now in the markets and stocks you care about". In practice, it is effectively used by finance professionals to express their opinions on individual firms and the market as a whole. Since mid-2010, users also have the possibility to label their own messages as either \textit{bullish} or \textit{bearish}. This feature is very useful for researchers as even though many messages are unlabeled, it allows for sentiment classification using NLP techniques. The reasons of using StockTwits and not another social media is threefold. First, one of the main challenge in Natural Language Processing is the creation of an appropriate labeled vocabulary. \citet{loughran2011barron}) shows that it is essential to have a specific vocabulary to interpret finance documents (i.e many words have a different meaning in finance than in traditional English (e.g “bear trap”)). In addition to that, social media slang is an additional layer of language complexity. To this extent, the functionality to self-tag \textit{bullish} and \textit{bearish} messages that StockTwits implemented in 2012 is very valuable as it allows the creation of a specific labeled vocabulary out of labeled messages. We are not aware of any other social media platform in finance offering this functionality. Second, StockTwits is less noisy than Twitter because messages focus on finance and economics matters only. Third, extracting data out of StockTwits is easy because of its API. StockTwits' API is designed to query the database to download messages via JSON requests. Using a Python script, we extract all messages since the 10th of July 2009 until the 31st of March 2020 for a preset list of tickers coming from the Compustat database corresponding to all US and Canadian firms. This results in 90 million messages, which we download and store as JSON files. As one message may refer to several tickers, we consider a message with two or more tickers as one message for every ticker identified. We refer to the appendix for more information about this process.

 \begin{figure}
    \centering
    \includegraphics[width=0.75\textwidth]{tweet_vol_daily.JPG}
    \caption{Number of messages posted daily on StockTwits.}
    \label{fig:tweet_vol_daily}
\end{figure}

Every message comes with the following eight features: (1) the ticker discussed (2) the exact timestamp of the message, (3) a unique message identifier, (4) the body of the message, (5) the sentiment label (bearish, bullish, or none) entered by the user, (6) a unique identifier of the user who sent the message, (7) the number of messages published by the user who sent the message, and (8) the number of followers of the user who sent the message. Figure~\ref{screenshot} shows a screenshot of the StockTwits website as of 3rd March 2020, for a query on the AAPL\footnote{AAPL is the ticker for Apple.} ticker. The first message is labeled as bullish by the user ``satkaru", the two next are unlabeled messages that will be classified using a machine learning classifier, and the last message is labeled as bearish by the user ``Etrading''.

\begin{figure*}[h]
\centering
\includegraphics[width=0.75\textwidth]{stocktwits_screeshot_3march2020.JPG}
\caption{Screenshot of StockTwits as of 3rd March 2020. The first message is labeled as bullish by the user ``satkaru", the two next are unlabeled messages that will be classified using a machine learning classifier and the last message is labeled as bearish by the user ``Etrading''.}
\label{screenshot}
\end{figure*}

Figure \ref{categrepartition-preclassif} shows the distribution of user-labeled and unlabeled messages. Overall, around 30 million messages are user-labeled and 60 million messages are unlabeled. Among the user-labeled messages we find five times more bullish than bearish ones. This ratio indicates that investors are on average optimistic about the market, which is consistent with findings in the literature, e.g., \citet{renault2017intraday}. Such class imbalance is a well-known issue in machine learning classification, which we address below. We will classify the unlabeled messages using a machine learning algorithm trained on the set of user-labeled messages. Since not every message contains substantial information, we believe that the sentiment classification should not be a bullish/bearish dichotomy. Hence, we allow for a neutral class to account for messages that do not take a clear stand.

\begin{figure*}[h]
\centering
        \begin{subfigure}[b]{0.75\textwidth}   
            \centering 
            \includegraphics[width=\textwidth]{preclassif.JPG}
        \end{subfigure}
        \caption{Number of user-labeled messages in each category. There are around 60 million unlabeled messages that we classify using the 30 million user-labeled bullish and bearish messages.} 
        \label{categrepartition-preclassif}
\end{figure*}

Figure \ref{fig:data-stocktwits1} shows a log-log histogram of the number of followers per user and a log-log histogram of the number of messages posted by users. As expected, there are few users with many followers (they can be seen as ``influencers"), and many users with a few followers. In addition, most users seem to post on average between 10 and 400 messages and a few post a lot more.

\begin{figure*}[]
\centering
        \begin{subfigure}{0.75\textwidth}   
            \centering 
            \includegraphics[width=\textwidth]{Followers.JPG}
        \end{subfigure}
        \quad
        \begin{subfigure}{0.75\textwidth}
            \centering
            \includegraphics[width=\textwidth]{Tweets.jpg}
        \end{subfigure}

        \caption{User summary statistics. Top graph is a log-log histogram of the number of followers per user and the bottom graph shows the log-log histogram of the number of messages posted by users.} 
        \label{fig:data-stocktwits1}
\end{figure*}

Figure \ref{fig:data-stocktwits2} shows the top 30 most discussed tickers on the platform. Of all tickers extracted, around 75\% are ordinary common share, 15\% are ETFs and the remaining 10\% are other types of securities. Not surprisingly, the S\&P index is the most discussed, followed by Apple and other big tickers. The messages about the 15 (30) biggest tickers represent 20\% (25\%) of the total number of messages, which indicates that users talk about a wide panel of tickers and not only big firms. The bottom graph shows a histogram of the number of messages per ticker. The x-axis is log-scaled because of extreme values, the distribution is highly skewed.

\begin{figure*}[]
\centering
\begin{subfigure}{0.75\textwidth}  
            \centering 
            \includegraphics[width=\textwidth]{top30tickers.jpg}
        \end{subfigure}
        \quad
        \begin{subfigure}{0.75\textwidth}   
            \centering 
            \includegraphics[width=\textwidth]{LogTweetsPerFirm_3.jpg}
        \end{subfigure}
        \caption{Firm summary statistics. Top graph shows the top 30 most discussed tickers on the platform. SPY is the ticker of the S\&P and AAPL is the ticker for Apple. Bottom graph shows the distribution of the number of messages across tickers.} 
        \label{fig:data-stocktwits2}
\end{figure*}

Messages contain qualitative information, which needs to be transformed into quantitative data for the computer to understand. Thereto, we first apply the following preprocessing operations: an apostrophe handler, a contraction form handler (i.e. ``aren't" becomes ``are not"), tickers removal, stop words (i.e. ``a", ``the", ``of") removal\footnote{We follow \citet{renault2020sentiment} and \citet{saif2014stopwords} and use a restrictive list of stopwords to avoid accuracy decrease.}, users removal, lemmatization, URLs removal and a simple spell corrector dealing with more than two repeated characters (i.e. ``soooo goooood" becomes ``soo good"). Table \ref{tab:preprocess} shows five examples of messages before and after preprocessing.

One of the first steps in NLP is the tokenization : the way of slicing a piece of text in smaller units called tokens or terms. In financial lingo, some words only have meaning when associated with other words (i.e ``bad apple" or ``bear flag"). N-gram models allow accounting for words frequently occurring together with other words. The main hyperparameter in N-gram models is the size of the group of words considered : unigram is a term with only one word, bigram is a term with two consecutive words, etc. Bigger N-grams models increase dramatically the size of the vocabulary (i.e : the collection of all terms considered). We are able to control the size of the vocabulary by tuning a hyperparameter keeping only a given number of most frequent terms.  We chose to work with a vocabulary consisting of the one million most frequent unigrams, bigrams or trigrams.

Figure \ref{wordclouds} represents the bullish and bearish word clouds. These correspond to the most frequent terms (up to 3-grams) in user-labeled bullish (bearish) messages relative to their total appearance. The size of the terms represent their importance in the cloud. In the bullish cloud, we see words such as ``bullish divergence", ``room to grow", ``lot potential" which are clearly bullish signals. In the bearish cloud, we find words such as ``recent resistance", ``short setup", ``bad apple" which indeed indicate bearish signals. These findings are reassuring in the sense that the content of the messages on the platform are consistent with their labels. The term ``aldox" in the bullish cloud caught our attention. After some research, it is an abbreviation for Aldoxorubicin, a drug against tumors and is associated with pharmaceutical messages where investors were very enthusiastic about it (example of related message: ``aldox is on the slide. have great faith this is truly world change"). That is why the term is appearing almost exclusively in bullish messages, hence in the bullish cloud.

On another note, most of the times messages are written by humans. It is however possible that some messages are generated by a robot (for instance to spread news or articles). In such cases, we define a neutral class that would absorb these messages if no substantial information is embedded in these messages. The term "long position open" present in the bearish cloud is an anomaly due to bearish user-labeled messages of intraday alerts such as “sell \$labd close labd long position. open labd short position. time: 14:53 ny price: \$13.64 zquant intraday alerts”. However, such anomalies are not an issue, a message “long position open” has a bullish probability of 0.91 and gets classify as bullish as it should. This is the case because when classifying a message, the score of the unigram “long” is much stronger than the score of the trigram “long position open”. From a linguistic point of view our approach is brute force. However, in turn, it works at a massive scale. Overall, the data quality is very good.

\begin{figure}
\begin{center}
    \includegraphics[width=0.75\textwidth]{Positive rate words.jpg}

    \vspace{1cm}

    \includegraphics[width=0.75\textwidth]{Negative rate words.jpg}
   \vspace{0.1cm}
    \caption{Bullish word cloud (top), bearish word cloud (bottom). These correspond to the most frequent terms (up to 3-grams) in user-labeled bullish (bearish) messages relative to their total appearance. The size of the terms represent their importance in the cloud.}
    \label{wordclouds}
  \end{center}
\end{figure}

\newpage
\begin{table}[H]
\begin{tabular*}{\textwidth}{l@{\extracolsep{\fill}}ll}
\hline \multicolumn{2}{l}{Before preprocessing}                         \\   \hline \hline                                                                          
(1) & @CassandraTwit \$uvxy contango 3.5\%...still long. goooooood      \\
(2) & \$FRPT Take profits while you still can. \\
(3) &  \$UVXY \$tvix go time boys and girls. Holding overnight again                                                                                                                                        \\
 (4) & \$dnr Nice upgrade as company goes into its quiet period!      \\
 (5) & \$SPY market won't reverse again towards closing. Get put options. \\ 
 \hline                                                                                          
\end{tabular*}
\end{table}

>>>>

\section{Introduction}
\label{S:3.1}

Can the stock market be predicted by analyzing social media? Recent developments in machine learning and the growing quantities of available text data from online news, social media and annual reports have triggered intensive research in finance. In their pioneering paper, \citet{antweiler2004all} compute a bullishness measure out of 1.5 million messages posted on Yahoo! Finance and Raging Bull and finds that stock messages help predict market volatility. Their results clearly reject the hypothesis that all that talk is just noise. They show that there is financially relevant information present in social media. In a similar vein, \citet{tet_07} constructs a measure of media pessimism from a Wall Street Journal column and finds that it predicts downward pressure on market prices.

Most of the previous financial studies of social media rely on pre-defined or manually annotated sentiment dictionaries. Such approaches are limited in various ways. How to create a sentiment classifier that understands the vocabulary of the messages posted by the investors? For instance, ``bull" is an animal in everyday language but it is someone optimistic in the financial jargon. \citet{loughran2011liability} create a word list, which helps classify tone in a financial document. However, this might not be sufficient in the context of social media because messages posted present many typos, abbreviations and slang, so one needs to have an additional layer of data preprocessing. For instance, the word ``goooooood" would not be recognized by the model if it is not corrected into ``good" first. On the other hand, manually annotating and validating dictionaries is not a scalable approach to handling social media content.

Our paper overcomes these limitations. We develop a machine learning algorithm to classify the sentiment of a large sample of StockTwits messages as bullish, bearish, or neutral. The sample consists of all messages referring to US and Canadian stocks, including ETFs and other types of securities available on CRSP/Compustat, from January 2010 to March 2020. We train our machine learning classifier on the set of all user sentiment-labeled messages, which constitute about one third of the sample. We then classify the sentiment of all remaining messages. Our method scales and performs very well. It achieves an out-of-sample accuracy of 85.9\%, which compares well to the anecdotal 80 to 85\% probability that human annotators agree on the sentiment of a document, see, e.g., \citet{wilson2005recognizing} and \citet{chen2020large}. As a side product, we generate a vocabulary of one million investor sentiment-labeled terms consisting of up to three words that frequently appear in StockTwits messages.

We then construct a stock-aggregate daily sentiment polarity measure and relate it to daily stock returns. We find that polarity is positively associated with contemporaneous returns. However, unconditionally, polarity cannot predict next-day returns, which is in line with the efficient market hypothesis (EMH). We then conduct an event study. We define events as days of sudden peaks of message volume of individual tickers. We classify events as bullish, bearish, or neutral depending on the prevailing polarities. We find that bullish (bearish) events are strongly associated with large positive (negative) abnormal returns. Cumulative abnormal returns over the preceding 20 days of an event have no predictive power on the type of event. Returns normalize immediately after the jump on the event date, which again is in line with the EMH. In contrast, remarkably, we find that cumulative abnormal polarity has statistically significant predictive power on the type of event. We assess the economic relevance of our findings with the performance of cumulative abnormal polarity ranked portfolios. We find that for appropriate choices of thresholds, cumulative abnormal polarities provide valuable signals for stock market investments.

As a technical byproduct, we develop a sentiment classifier of micro-blogs for imbalanced data. This addresses the stylized fact that bloggers post more bullish than bearish-labeled messages. In our sample, the ratio is five to one. On the other hand, we find that not all messages carry a substantial stock market relevant sentiment. Rather than re-sampling from the underrepresented bearish class, we thus introduce an auxiliary neutral class. We then run two independent binary classifiers. The first (second) classifies messages as bullish versus non-bullish (bearish versus non-bearish). We aggregate the two binary outcomes and classify a message as bullish (bearish) for the concordant combination bullish/non-bearish (non-bullish/bearish), and neutral otherwise. This approach is very simple and efficient, and eliminates the class imbalance bias at the same time. It builds on any traditional binary classifiers. We use logistic regression on Term Frequency-Inverse Document Frequency (TFIDF)-vectorized messages. TFIDF is a weighting scheme gauging the importance of a word in a document.

Our paper contributes to the growing literature on machine learning classification of social media and its interaction with the stock market. Most previous financial studies use Twitter as their primary source of data. Twitter has the advantage of being used by a wide range of people across the world and a few influencers can attract the attention of many investors. In 2013, following a meeting with Tim Cook (Apple CEO), Carl Icahn tweeted that he bought a large position in Apple and believed that the company is extremely undervalued. This bullish tweet caused the market capitalization of Apple to jump by \$12 billion. In 2019, JPMorgan has created the Volfefe Index to track Donald Trump's tweets impact on the stock market. However, it is more difficult to disentangle noise from relevant tweets in Twitter than in other more focused social media. Results from \citet{Ghoshal} show that StockTwits is significantly more informative than Twitter data. This is not surprising as StockTwits is a finance-focused platform whereas Twitter also captures irrelevant opinions on a wide range of non-finance related matters.

Our paper is the first that analyzes the predictive power of StockTwits messages on stock returns unconditionally and around specific events. \citet{renault2017intraday} builds an intraday investor sentiment indicator using messages and finds that the change in investor sentiment of the first half-hour of a trading day helps forecast the last half-hour market return of that trading day. However, his classifier is based on a dictionary consisting of 8 thousand manually validated and modified terms, which limits its scalability. \citet{renault2020sentiment} uses larger data sets and compares various classifiers, including machine learning.

Our approach is in some parts similar to \citet{ranco2015effects}, who also study the relation of micro-blog sentiments with stock returns. However, they use Twitter data, whereas the finance-tailored StockTwits data we use results in higher contemporaneous correlations between stock returns and polarity. They manually annotate 100 thousand tweets, which limits the scalability of their approach. Our sample is much larger (90 million versus 1 million messages) and covers a longer period (10 years versus 13 months).

\citet{ke_kel_xiu_20} extract sentiment from news articles on the Dow Jones Newswires. They train a  sentiment score directly on returns. In contrast, we use user sentiment-labeled StockTwits messages as training and validation set for our sentiment classifier.

Our paper also contributes to the EMH literature by gauging how cumulative average abnormal returns and abnormal polarities behave around sudden peaks of message activity.

The remainder of the paper is structured as follows. Section \ref{S:data} discusses StockTwits and stock market data. Section \ref{S:classification} develops our sentiment classifier based on TFIDF vectorization. Section \ref{S:polarity} introduces the sentiment polarity measure and relates it to stock returns. Section \ref{S:event studies} contains the event study. Section \ref{S:portfolios} discusses the sentiment-sorted portfolio performance. Section \ref{S:conclusion} concludes. The appendix contains additional statistics and background material.

\section{StockTwits and Stock Market Data}
\label{S:data}

StockTwits is a large social media platform similar to Twitter but designed for investors and traders. Users register online and can post messages about any listed stock through the prefix \$ followed by the ticker of the stock. StockTwits was created in 2008 as an app built on the Twitter's API and later detached from Twitter to build a standalone social network. As of April 2019, it had over two million registered users and the number of daily posted messages has been growing exponentially, see Figure~\ref{fig:tweet_vol_daily}.

\clearpage

\begin{figure}[h]
    \centering
    \includegraphics[width=0.75\textwidth]{tweet_vol_daily.JPG}
    \mycaption{Number of messages posted daily on StockTwits}{Message volume posted on StockTwits. Numbers are aggregated daily.}
    \label{fig:tweet_vol_daily}
\end{figure}

StockTwits describes itself as ``the voice of social finance and the best way to find out what is happening right now in the markets and stocks you care about''. In practice, it is effectively used by finance professionals to express their opinions on individual stocks and the market as a whole. Importantly, users have the option to label their posted messages as either bullish or bearish.\footnote{This optional label was effectively available as of mid-2010.} This feature is key for our approach, as it allows for sentiment classification of all messages using machine learning trained on the user-labeled messages.\footnote{One third of the messages in our sample have a user labeled sentiment.}

The reasons for using StockTwits and not other social media data for financial studies are at least threefold. First, a major challenge in applying natural language processing is the creation of an appropriate labeled vocabulary. \citet{loughran2011barron} show that it is essential to have a specific vocabulary to interpret finance documents (i.e., many words have a different meaning in finance than in traditional English, such as “bear trap”). In addition to that, social media slang is an additional layer of language complexity. To this extent, the functionality to self-tag bullish and bearish messages that StockTwits is extremely valuable as it allows the creation of a specific labeled vocabulary out of labeled messages. We are not aware of any other social media platform in finance offering this functionality. Second, text data from StockTwits is more reliable and less noisy than from general purpose platforms, such as Twitter, because messages focus on finance and economics matters only. Micro-bloggers have incentives to post valuable information in order to maintain or increase mentions and retweets, and thus have a greater share of voice in the forum (\citet{sprenger2014tweets}). On the other hand, StockTwits messages might be biased and subject to malicious users that try to manipulate the market. However, market manipulations likely happen only rarely as the SEC closely monitors potential influencers to prevent any market abuse. Third, extracting data from StockTwits is easy because of its API. StockTwits' API is designed to query the database to download messages via JSON requests. We provide a short tutorial in Appendix~\ref{app_tut}.

We use stock market data from CRSP/Compustat. We extract daily closing prices, daily volume of transactions and number of shares outstanding for all US and Canadian stocks, as well as ETFs and some other types of securities, from January 2010 to March 2020. Stock prices and number of shares are adjusted to account for any distribution (i.e., dividends, stock splits) so that a comparison can be made on an equivalent basis before and after the distribution. We use as risk-free rate the 3-month US T-bill rate, converted into daily risk-free returns. We henceforth refer to daily stock excess returns over risk-free simply as returns. Using a Python script, we then extract all messages from StockTwits for the list of tickers corresponding to the sample of US and Canadian stocks. This results in 90 million messages, which we download and store as JSON files.\footnote{A message may refer to multiple tickers. We count any such message towards any ticker that it refers to. We give more information about this double counting in Appendix \ref{app_counting}.} Overall, our sample covers 8843 tickers, whereof 75\% refer to ordinary common share, 15\% to ETFs, and the remaining 10\% to other types of securities. Henceforth, we interchangeably refer to any of these securities as either a stock or a ticker.

Every StockTwits message includes eight features: (1) the reference ticker(s), (2) a timestamp, (3) a unique message identifier, (4) the body of the message, (5) the sentiment label (bearish, bullish, or none) entered by the user, (6) a unique identifier of the user who posted the message, (7) the number of messages published by the user who posted the message, and (8) the number of followers of the user who posted the message. Our sentiment analyses builds on the first five features. The last three provide additional information on the network structure, which we briefly discuss in Appendix \ref{app_userstats}.

Figure~\ref{screenshot} shows a screenshot of the StockTwits website as of 3rd March 2020, for a query of AAPL, which is the ticker for Apple. The first message is labeled as bullish by the user ``satkaru", the two next are unlabeled messages that will be classified by our machine learning algorithm, and the last message is labeled as bearish by the user ``Etrading''.

\begin{figure}[h]
\centering
\includegraphics[width=0.75\textwidth]{stocktwits_screeshot_3march2020.JPG}
\mycaption{Screenshot of messages posted on StockTwits}{Screenshot is from the 3rd of March 2020 for a query on AAPL, the ticker of Apple.}
\label{screenshot}
\end{figure}

Left plot of Figure \ref{fig:data-stocktwits2} shows the top 30 most discussed tickers on StockTwits. SPY, a large ETF that tracks the S\&P 500 stock market index, is the most discussed ticker, followed by Apple and other big tickers. The messages about the 15 (30) biggest tickers represent 20\% (25\%) of the total number of messages, which indicates that users talk about a wide panel of tickers and not only big firms. Right graph shows a histogram of the number of messages per ticker. The x-axis is log-scaled because due to extreme values the distribution is highly skewed.

\begin{figure*}[h]
\centering
\begin{subfigure}{0.49\textwidth}  
            \centering 
            \includegraphics[width=\textwidth]{top30tickers.jpg}
        \end{subfigure}
        \begin{subfigure}{0.49\textwidth}   
            \centering 
            \includegraphics[width=\textwidth]{LogTweetsPerFirm_3.jpg}
        \end{subfigure}
        \mycaption{Ticker summary statistics}{
        Left graph shows the top 30 most discussed tickers on StockTwits. SPY is the ticker of a large ETF tracking the S\&P 500 and AAPL is the ticker for Apple. Right graph shows the distribution of the number of messages across tickers.} 
        \label{fig:data-stocktwits2}
\end{figure*}

Text messages need to be transformed into a quantitative vector to be fed into our machine learning classifier, which in turn computes a sentiment score. This transformation consists of several steps. First, we apply some preprocessing operations to the text messages: an apostrophe handler, a contraction form handler (e.g., ``aren't" becomes ``are not"), tickers removal, stop words removal (e.g., ``a", ``the", ``of")\footnote{We follow \citet{renault2020sentiment} and \citet{saif2014stopwords} and use a restrictive list of stopwords to avoid accuracy decrease.}, users removal, lemmatization, URLs removal and a simple spell corrector dealing with more than two repeated characters (e.g., ``soooo goooood" becomes ``soo good"). Table \ref{tab:preprocess} shows five examples of messages before and after preprocessing.

\begin{table}[h]
\centering
\begin{tabular}{ll} \hline
\multicolumn{2}{l}{Before preprocessing}                                   \\ \hline  
(1) & @CassandraTwit \$uvxy contango 3.5\%...still long. goooooood        \\
(2) & \$FRPT Take profits while you still can.                            \\
(3) & \$UVXY \$tvix go time boys and girls. Holding overnight again       \\
(4) & \$dnr Nice upgrade as company goes into its quiet period!           \\
(5) & \$SPY market won't reverse again towards closing. Get put options.  \\ \hline   \hline
\multicolumn{2}{l}{After preprocessing}                                   \\ \hline 
(1) & contango still long good                                            \\
(2) & take profit while you still can                                     \\
(3) & go time boy and girl hold overnight again                           \\
(4) & nice upgrade as company go into its quiet period                    \\
(5) & market will not reverse again towards closing get put options      \\ \hline
\end{tabular}
\mycaption{Preprocessing of five sample messages}{Preprocessing operations include: punctuation removal, lower casing, apostrophe handling, contraction form handling (i.e., ``won't" becomes ``will not"), tickers removal, users removal, URLs removal, parsing and a simple spell corrector dealing with more than two repeated characters (i.e., ``goooood" becomes ``good")}
\label{tab:preprocess}
\end{table}

The next step is tokenization: the slicing of a text message into smaller units called terms or tokens. In financial lingo, some words only have meaning when associated with other words (i.e., ``bad apple" or ``bear flag"). $N$-gram models allow accounting for words frequently occurring together with other words. The main hyperparameter in an $N$-gram model is the number $N$ of words that form a term: a unigram is a term with only one word, a bigram is a term with two consecutive words, etc. Larger $N$-gram models increase dramatically the size of the vocabulary (i.e., the collection of all terms considered). We select $N=1,2,3$ and truncate the resulting vocabulary such that it consists of the one million most frequent terms in the union of all unigrams, bigrams and trigrams.

\begin{figure}[h]
\centering
\begin{subfigure}{0.49\textwidth}
\centering
\includegraphics[width=1\textwidth]{Positive rate words.jpg}
\end{subfigure}
\begin{subfigure}{0.49\textwidth}
\centering
        \includegraphics[width=1\textwidth]{Negative rate words.jpg}
\end{subfigure}
    \mycaption{Word clouds}{Bullish word cloud (left), bearish word cloud (right). These correspond to the most frequent terms (up to trigrams) in user-labeled bullish (bearish) messages relative to their total appearance. The size of the terms represents their importance in the cloud.}
    \label{wordclouds}
\end{figure}

Figure \ref{wordclouds} represents the bullish and bearish word clouds. These represent the most frequent terms in all user-labeled bullish and bearish messages relative to their total appearance, respectively. The size of the terms represents their relative weight in the cloud. In the bullish cloud, we see terms such as ``bullish divergence", ``room to grow", ``lot potential" which we can clearly interpret as bullish signals. In the bearish cloud, we find terms such as ``recent resistance", ``short setup", ``bad apple" which again we can directly interpret as bearish signals. These findings are reassuring in the sense that the content of the messages on StockTwits are consistent with their labels. We checked for anomalies at random, but did not find significant issues. Appendix \ref{app_ano} discusses two such anomalies.

%\begin{figure}[h]
%\begin{center}
%    \includegraphics[width=0.5\textwidth]{Positive rate words.jpg}

%    \vspace{1cm}

%    \includegraphics[width=0.5\textwidth]{Negative rate words.jpg}
%   \vspace{0.1cm}
%    \caption{Bullish word cloud (top), bearish word cloud (bottom). These correspond to the most frequent terms (up to trigrams) in user-labeled bullish (bearish) messages relative to their total appearance. The size of the terms represent their importance in the cloud.}
%    \label{wordclouds}
%  \end{center}
%\end{figure}

\section{Sentiment Classification}\label{S:classification}

Left plot of Figure \ref{fig:classif_time} shows the proportions of user sentiment-labeled messages across time. In the early years of the platform, most messages were unlabeled, presumably because users were not familiar with the sentiment label yet. Albeit the proportion of unlabeled messages monotonically declines over the years, almost 60\% of the more recent messages are still unlabeled. Overall, around 30 million messages are user-labeled and 60 million messages are unlabeled. We conjecture that by far not all unlabeled messages reflect market neutral opinions. Indeed, the right plot of Figure \ref{fig:classif_time} reveals that a substantial part of user-unlabeled messages is machine learning classified as bullish or bearish. Hence these user-unlabeled messages contain indeed market relevant information, which we are able to capture by our algorithm.

%\begin{figure}[h]
%    \centering
%    \includegraphics[width=0.5\textwidth]{preclassif_acrosstime.JPG}
%    \caption{Proportions of user-labeled messages: bullish (green), bearish (red), and unlabeled (gray) across time. Proportions are aggregated monthly.}
%    \label{fig:ts_neutral_vs_uncl}
%\end{figure}

\begin{figure}[h]
\centering
\begin{subfigure}{0.49\textwidth}
\centering
\includegraphics[width=1\textwidth]{preclassif_acrosstime.JPG}
\end{subfigure}
\begin{subfigure}{0.49\textwidth}
\centering
        \includegraphics[width=1\textwidth]{postclassif_acrosstime3.JPG}
\end{subfigure}

    \mycaption{Message classification}{Left plot shows the proportions of user-labeled messages: bullish (green), bearish (red), and unlabeled (gray) across time. Right plot shows the proportions of machine learning classified messages: bullish (light green predicted, green user-labeled), bearish (light red predicted, red user-labeled), and neutral (gray) across time. Proportions are aggregated monthly.}
    \label{fig:classif_time}

\end{figure}
